---
title: "Maximum likelihood estimator"
author: "Manuel Arce"
date: "2024-01-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Delaporte)
```

# Using seed 123, generate a sample of size 20 from a negative binomial distribution with size = 20 and prob = 0.3

```{r}
set.seed(123)
sample_data <- rnbinom(n = 20, size = 20, prob = 0.3)
sample_data
```

# Fit a geometric distribution to the generated sample from exercise 3. What is the estimated value of p?

# Maximum likelihood estimator

```{r}
p_mle <- 1 / (mean(sample_data) + 1)
p_mle
```

# Maximum likelihood estimator using the optim function

```{r}
geom_model <- optim(par = p_mle,
                    fn = function(par) {
                        -sum(
                            log(
                                dgeom(x = sample_data, prob = par)
                            )
                        )
                    },
                    method = "L-BFGS-B",
                    lower = 0.0001,
                    upper = 0.9999)
geom_model$par
```

# Fit a Delaporte distribution by maximum likelihood using the generated sample.

# What values do you estimate for α, β, λ?

```{r}

beta <- ((mean((sample_data - mean(sample_data))^3) - var(sample_data)) /
         (2 * (var(sample_data) - mean(sample_data)))) - 1
alpha <- (var(sample_data) - mean(sample_data)) / beta^2
lambda <- mean(sample_data) - alpha * beta

delaporte_model <- optim(par = c(alpha, beta, lambda),
                         fn = function(par) {
                             -sum(
                                 log(
                                     ddelap(x = sample_data,
                                            alpha = par[1],
                                            beta = par[2],
                                            lambda = par[3])
                                 )
                             )
                         },
                         method = "L-BFGS-B",
                         lower = c(0.01, 0.01, 0.01),
                         upper = c(70, 70, 80))
delaporte_model$par
```
